---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r unzip, cache = TRUE, echo = TRUE}
library(ggplot2)
library(plyr)
unzip("activity.zip",exdir = "data")

```

```{r reading, cache = TRUE, echo = TRUE}

activity <- read.csv("data/activity.csv", stringsAsFactors=FALSE)
str(activity)
summary(activity)
```


##pulling data without nas
```{r, echo = TRUE}
activity$day <- weekdays(as.Date(activity$date))
activity$DateTime<- as.POSIXct(activity$date, format="%Y-%m-%d")

##pulling data without nas
clean <- activity[!is.na(activity$steps),]
head(activity$date)
head(activity$day)
```
## What is mean total number of steps taken per day?
```{r, echo = TRUE}
TotalStepsPerDay <- tapply(activity$steps, activity$date, sum)
hist(TotalStepsPerDay, xlab = "Number of Steps", main = "Histogram: Steps per Day")

```
###Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = TRUE}
meanStepsPerDay <-  mean(TotalStepsPerDay, na.rm = TRUE)
medianStepsPerDay <-  median(TotalStepsPerDay, na.rm = TRUE)
```


## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo = TRUE}
StepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(as.numeric(names(StepsPerInterval)), 
     StepsPerInterval, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Average Daily Activity Pattern", 
     type = "l")
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo = TRUE}
maxInterval <- names(sort(StepsPerInterval, decreasing = TRUE)[1])
maxSteps <- sort(StepsPerInterval, decreasing = TRUE)[1]
```
## Imputing missing values
```{r, echo = TRUE}
```


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE}
```